Transient Corotating Clumps Around Adolescent Low-Mass Stars From Four Years of TESS

### 2023-08-31 paper/ms.tex
update ms

Wrapping is necessary to eliminate edge effects when fitting the light
curve and when identifying local minima in the residuals.  A phased
light curve would usually cover phases $\phi \in [ 0,1 ]$.  We instead
performed the analysis described above using a phase-folded light
curve spanning $\phi \in [-1,2 ]$, which was created by duplicating
and concatenating the ordinary phase-folded light curve.  The free
parameters we adopted throughout the analysis -- for instance the
maximum number of spline knots per cycle, and the height and depth
criteria for dips -- were chosen during testing based on the desire to
correctly re-identify a large fraction ($>$90\%) of previously known
CRs, while also being able to consistently reject common false
positives such as rapidly rotating spot-induced variability and
typical eclipsing binaries.

@@

Referring back to Figure~\ref{fig:catalogscatter}, which compares the
properties of CRs with those of the search sample (top panels) and the
Pleiades \citep{2016AJ....152..114R}, we see that the typical CR
masses span 0.1-0.4\,$M_\odot$, and the typical ages span 2-150\,Myr.
This mass and age range includes both fully convective stars and stars
with a combination of radiative cores and convective envelopes, with
the dividing line at around $M_\star = 0.25\,M_\odot$
\citep{2018A&A...619A.177B}.  We found no obvious differences in light
curve morphology between CRs above and below the boundary of
0.25\,$M_\odot$.  In terms of their rotation rates relative to the
Pleiades, the CRs are among the more rapidly rotating half of M
dwarfs.

@@

In Section~\ref{subsec:binarity}, we found that a significant fraction
of the CRs show indications of unresolved binarity.  Excess noise
above the Gaia single-source astrometric model is common
(\ngoodhighruwe/\ngoods\ high-quality CRs with RUWE$_{\rm DR3}$$>$2),
as is the presence of multiple periods in the TESS light curves
(\ngoodmultperiodflag/\ngoods).  Elevated astrometric noise is almost
always accompanied by multiple detectable TESS periods
(\ngoodruweandmultperiod/\ngoodhighruwe\ high-quality CRs).  The
latter observation corroborates the claim that most sources with
RUWE$_{\rm DR3}$$>$2 are binaries with projected apparent separations
below 1$''$, and projected physical separations $\lesssim$50\,AU.

One possible relation between binarity and CRs is that the presence of
an intermediate-separation binary could cause early disk dispersal,
freeing the star to contract.  The CR phenomenon seems to require
rapid rotation; thus if binary systems tend to produce more rapidly
rotating stars, we would expect a sample of CRs to have a larger
binary fraction than the field.  We do not see clear evidence for such
an effect.  The multiplicity rate of M dwarfs near the Sun is
26.8$\pm$1.4\% \citep{2019AJ....157..216W}.  Based on the same study,
the peak of the separation distribution decreases from $\approx$49\,AU
for $0.30$-$0.60$\,$M_\odot$ stars, to $\approx$11\,AU for
$0.15$-$0.30$\,$M_\odot$ M dwarfs.  The multiplicity fraction in our
CR sample seems approximately consistent with this field fraction.

@@

We originally classified TIC~435903839, with RUWE$_{\rm DR3}$=17.7, as
an ``ambiguous'' CR with a 10.8\,hr period, because this period
minimized the dispersion in the phase-folded light curve.  More
careful inspection revealed an impostor:
this source is a photometric blend of two ordinary rotating stars with
$P_0$=3.60\,hr, and $P_1$=5.41\,hr, giving a beat period $(P_0^{-1} -
P_1^{-1})^{-1}$ of 10.8\,hr.  This is a novel false positive scenario
for CRs: two rapidly-rotating stars with near-integer ratios of
rotation periods.  The beat between the two rotation signals produces
the apparent CR signal.  Such false positives can be excluded through
careful accounting of all peaks in a periodogram.  For instance,
TIC~435903839 shows a peak at 16.27\,hr, which is not an integer
multiple of the dispersion-minimizing 10.82\,hr period.

@@

While CR periods appear to remain constant to within measurement
precision over thousands of cycles, the light curve shapes evolve over
typical timescales of 10 to 1{,}000 cycles (e.g.
Figures~\ref{fig:evoln} and~\ref{fig:lpriver0}).  The CRs are
therefore quasiperiodic, with a characteristic dip longevity of
$\approx$100 cycles.  While this observation is consistent with the
analyses by \citet{2022AJ....163..144G} and
\citet{2023ApJ...945..114P}, it marks a qualitative departure from the
``persistent'' {\it vs.} ``transient'' flux dip distinction described
by \citet{2017AJ....153..152S}, since all CR dips seem to be transient
over timescales of more than 1{,}000 cycles (Figure~\ref{fig:evoln}).

@@

\subsubsection{What is a prominence?}
The prominence idea is based on an analogy with quiescent
prominences/filaments in the solar corona that last as long as a few
weeks \citep[see][]{2015ASSL..415.....V}.  In the context of the Sun,
a prominence is a clump of cold, partially ionized hydrogen viewed in
emission against the dark backdrop of space.  A filament is the same
clump of plasma, but viewed in absorption against the solar disk.  In
an extrasolar context, spectroscopic detections of transient Balmer-
and resonance-line absorption seen for stars such as AB~Dor and
Speedy~Mic
\citep[e.g.][]{1989MNRAS.238..657C,1993MNRAS.262..369J,2006MNRAS.365..530D,2016MNRAS.463..965L}
have been interpreted as prominences that scatter a star's
chromospheric emission \citep[see][]{1989MNRAS.238..657C}.  The
short-term mechanical stability of such gas configurations is
theoretically plausible for rapidly rotating stars
\citep{2000MNRAS.316..647F,2022MNRAS.514.5465W}.   To our best
knowledge, this class of spectroscopic observation also has no viable
alternative explanations.

We performed a simple visual examination of the TESS light curves for
five prominence-hosting systems studied by
\citet{2019MNRAS.482.2853J}--AB~Dor, Speedy~Mic, LQ Lup, HK Aqr, and
V374 Peg--and detected no CR behavior.  Assuming that
spectroscopically observable prominence systems do not turn off, this would
imply that they are not always accompanied by photometric CR-like
dips: a link between the spectroscopic prominences
that may exist around rapidly rotating low-mass stars and the CR
phenomenon has yet to be made.

@@

CRs show broadband flux variations that can be 1-2$\times$ deeper in
the blue than in the red
\citep{2017PASJ...69L...2O,2020AJ....160...86B,2022AJ....163..144G,2023MNRAS.518.2921K}.
Dust can naturally explain this chromaticity, since it has a larger
absorption cross-section in the blue than the red
\citep[e.g.][]{1989ApJ...345..245C}.  Gas might also explain the observed
chromaticities \citep{1992oasp.book.....G}.  While bound-bound
absorption can be excluded, since it provides opacity only at narrow
resonant lines, the (neutral) hydrogen opacity due to bound-free
absorption is ``jagged'' \citep[see][Figure 8.5 and
Eq.~8.8]{1992oasp.book.....G}, such that at temperatures of
$\approx$3{,}000\,K to $\approx$10{,}000\,K the opacity can be larger
at blue wavelengths than at red wavelengths.  Bound-free absorption of
${\rm H}^{-}$ is often important at such temperatures, but this
opacity source is stronger in the red than the blue, the opposite of
what is required to produce deeper dips in the blue than in the red.
Likewise, Thomson scattering is too gray to be the dominant opacity
source.

@@

The observed lifetime of the CR phenomenon could provide another way
to discern between the gas and dust clump scenarios.  Based on the
available statistics from e.g.  \citet{2022AJ....164...80R} and
references therein, it seems plausible that CR occurrence decreases
with stellar age from $\approx$3\% at 10\,Myr (Sco-Cen),
to $\approx$1\% at 100\,Myr (Pleiades),
down to 0\% by the $\approx$700\,Myr
age of Praesepe.  This is odd in the context of the prominence
scenario, because pre-main-sequence M dwarfs spin up over the first
100\,Myr; prominences might therefore even be expected to be {\it
more} common at 100\,Myr than at 10\,Myr, under the assumption that the
production of prominences depends only on the stellar rotation rate. 
The dust clump scenario would hold a natural explanation: the lower occurrence of CRs around older stars would
simply reflect a finite supply of dust.
One possible complication however is that the magnetic field topology
of rapidly rotating M dwarfs may depend on factors other than the
rotation rate (e.g. the age), which might affect the production of prominences.

@@

In the context of disk-driven planet migration, the stopping location
for the innermost planet is set by the location of the protoplanetary
disk's truncation radius \citep{2018haex.bookE.142I}.  The 
truncation radius is traditionally calculated by equating the magnetic
pressure from the stellar magnetosphere with the ram pressure of the
inflowing gas.  As it happens, the truncation radius is close to the
corotation radius for low accretion rates
\citep{2016ARA&A..54..135H,2022MNRAS.510.5246L}.  These considerations
invite us to imagine one or more planets migrating inward due to gas
drag, and arriving at $\approx$5-10 stellar
radii before the disk is depleted.

With this picture in mind, it is tempting to attribute features of the
CR light curves as transits of material ejected by planets or
planetesimals.  Young rocky bodies are expected to be hot, and they
might expel either gas or dust.  The Jupiter-Io
system \citep[e.g.][]{2004jpsm.book..537S} is analogous, in that a small rocky
body similarly feeds the construction of a plasma torus.  We emphasize
that while this type of configuration seems a priori plausible, no
direct evidence currently supports it.

The main logical function of the planetesimals would be to serve as a
source for the occulting gas or dust; they would not necessarily need
to explain the observed phases of the observed dips.  The azimuthal
angle of the eventual entrainment would be entirely determined by the
stellar magnetic field.  In this scenario, the obscuring material
would inspiral from one or more rocky bodies well beyond the
corotation radius.  The planetesimals themselves would not necessarily
need to transit.  However if they did, they would need to be $\lesssim
1$\,$R_\oplus$ based on their non-detections in the TESS data.
Possibly analogous systems include K2-22 \citep{2015ApJ...812..112S}
and KOI-2700 \citep{2014ApJ...784...40R}, though the obscuring
material in the CRs would need to be observed much further from the
emitting planet than for those two examples.

@@

Although this outlier is intriguing, stars without infrared excessess
typically have more stable optical light curves than those with
infrared excesses.  While some dippers may evolve into CRs after the
disk is mostly gone, this would be generically expected based on the
population statistics: young stars become older.  There may be no
other causal connection between the two evolutionary stages.  With
that said, a common mystery between the CRs and dippers is how exactly
the {\it narrowness} of the dust-induced dimmings is produced.  It is
not unreasonable to imagine a similar mechanism operating for both
types of object, tied perhaps to a shared magnetic topology, or
perhaps to a preference for dust to inspiral to the star in clumped
structures.

@@

If the disappearance of a dip represents the permanent loss of the
obscuring material -- for example, if it is the result of a dust clump
being accreted or ejected -- then we can also estimate the rate at
which mass is flowing through the structures that lead to dips.  For
instance, LP~12-502 showed three ``state-switch'' events over the six
months of available TESS observations, during cycles 261, 309, and
1241 (Figure~\ref{fig:lpriver0}).  The other source for which we
performed a comparable analysis, TIC~300651846
(Appendix~\ref{app:tic3006}), showed two state-switches over 11
months.  In all such cases, at least one dip turned off.  For purposes
of estimation, we will take LP 12-502 as our prototype.  Assuming the
occulting material is dust, the corresponding $\dot{M} \equiv M\cdot
{\rm d}N/{\rm d}t$ time-averaged over six months is
$\approx$3$\times$$10^{-18} M_\odot\,{\rm yr}^{-1} \approx
1\times10^{-12} M_\oplus\,{\rm yr}^{-1}$.  Considered cumulatively
over the $\approx$$10^8$ years for which the CR phenomenon is
observed, this yields a cumulative moved dust mass of
$10^{-4}\,M_\oplus$, of order the Solar System's asteroid belt.  If
the occulting material is gas, the masses involved would be of order
100 times larger.  For cases in which we observe the {\it growth} of
dips, such as the Sector~29 data for TIC~224283342, or Sector~5 of
TIC~294328885, the dip depths typically increase by of order a few
percent over ten to twenty days.  This growth rate yields a mass flux
one order of magnitude larger than the earlier estimate.

@@

In this work, we searched 120-second cadence TESS data collected from
2018 July to 2022 September for complex rotators (CRs).  The target
stars were \nstarssearched\ K and M dwarfs within 150\,pc.  The
selection function was $>$$80\%$ complete within 30\,pc, and $<$$10\%$
complete at distances exceeding 100\,pc.

In our target sample, we found \ngoods\ objects that showed complex
quasiperiodic behavior over at least one TESS sector.  These \ngoods\
bona fide CRs are listed in Table~\ref{tab:thetable}.  This table also
includes \nmaybes\ ambiguous CRs, whose designation is
less certain, and \ndebunked\ impostors.  We inferred ages for all but
two of the \nallcands\ objects based on memberships in young stellar assocations; we
also derived temperatures and radii using SED fitting, and 
inferred stellar masses by interpolating against stellar evolutionary
models.  We caution that our sample is far from being
volume-limited and is not even magnitude-limited:
the TESS 120-second stellar sample had a heterogeneous
selection function which may have been biased in favor of young stars
over field stars.  Previous work however has shown that $\approx$1-3\%
of M dwarfs younger than $\approx$100\,Myr show the CR phenomenon
\citep{2016AJ....152..114R,2022AJ....163..144G}.

@@

      on a quasi-sinusoidal signal that presumably comes from
      starspots and faculae that produce inhomogeneities in the
      longitudinal flux distribution across the stellar surface.  The
      only viable explanation currently known for the sharp dips being superposed on
      the quasi-sinusoidal signals is that concentrations (``clumps'')
      of circumstellar material are corotating with the star.

@@

      the shortest dips, the transiting material must therefore also
      be capable of having effective sizes comparable to the star.

@@

%{\bf AUTHOR CONTRIBUTION STATEMENTS: PLEASE REVISE IF APPROPRIATE)}
LGB conceived the project and performed the dip-counting search, light curve classification, cluster membership, SED, variability, and secondary-period analyses.
RJ and SR performed the Fourier-based analysis and contributed to light curve classification.
LR cross-examined the light curve classification, and contributed an independent SED analysis.
AD identified the magnetic B star connection.
LAH contributed to project design and to the interpretation of the light curves.
JNW and SR significantly improved the clarity of the manuscript.
G\'AB acquired and maintained the servers used to run the dip-finding pipeline.
%-----
%{\bf JNW, MNG, GRR to confirm coauthorship \&/or to help out on interpretation!}
% Pending!
%HI and AH designed and maintain the architecture used to acquire and process the HIRES spectra.
All authors assisted in manuscript revision.

>>>>

Wrapping is necessary to eliminate edge effects when fitting the light
curve and when identifying local minima in the residuals.  A phased
light curve would usually cover phases $\phi \in [ 0,1 ]$.  We instead
performed the analysis described above using a phase-folded light
curve spanning $\phi \in [-1,2 ]$, which was created by duplicating
and concatenating the ordinary phase-folded light curve.  The free
parameters we adopted throughout the analysis -- for instance the
maximum number of spline knots per cycle, and the height and depth
criteria for dips -- were chosen during testing based on the desire to
correctly re-identify a large fraction ($>$90\%) of previously known
CRs, while also being able to consistently reject common false
positives such as spot-induced variability and
eclipsing binaries.

@@

Referring back to Figure~\ref{fig:catalogscatter}, which compares the
properties of CRs with those of the search sample (top panels) and the
Pleiades \citep{2016AJ....152..114R}, we see that the typical CR
masses span 0.1-0.4\,$M_\odot$, and the typical ages span 2-150\,Myr.
This mass and age range includes both fully convective stars and stars
with a combination of radiative cores and convective envelopes, with
the dividing line for these ages at around $M_\star = 0.25\,M_\odot$
\citep{2018A&A...619A.177B}.  We found no obvious differences in light
curve morphology between CRs above and below the boundary of
0.25\,$M_\odot$.  In terms of their rotation rates relative to the
Pleiades, the CRs are among the more rapidly rotating half of M
dwarfs.

@@

In Section~\ref{subsec:binarity}, we found that a significant fraction
of the CRs show indications of unresolved binarity.  Excess noise
above the Gaia single-source astrometric model is common
(\ngoodhighruwe/\ngoods\ high-quality CRs have RUWE$_{\rm DR3}$$>$2),
as is the presence of multiple periods in the TESS light curves
(\ngoodmultperiodflag/\ngoods).  Elevated astrometric noise almost
always implies multiple detectable TESS periods
(\ngoodruweandmultperiod/\ngoodhighruwe\ high-quality CRs).  The
latter observation corroborates the claim that most sources with
RUWE$_{\rm DR3}$$>$2 are binaries with projected apparent separations
below 1$''$, and projected physical separations $\lesssim$50\,AU.

One possible relation between binarity and CRs is that the presence of
an intermediate-separation binary could cause early disk dispersal,
freeing the star to contract.  The CR phenomenon seems to require
rapid rotation; thus if binary systems tend to produce more rapidly
rotating stars, we would expect a sample of CRs to have a larger
binary fraction than the field.  We see marginal, if any, evidence for such
an effect.  The multiplicity rate of M dwarfs near the Sun is
26.8$\pm$1.4\% \citep{2019AJ....157..216W}.  Based on the same study,
the peak of the separation distribution decreases from $\approx$49\,AU
for $0.30$-$0.60$\,$M_\odot$ stars, to $\approx$11\,AU for
$0.15$-$0.30$\,$M_\odot$ M dwarfs.  The multiplicity fraction in our
CR sample (32\% from RUWE$_{\rm DR3}$) seems approximately consistent
with this field fraction.

@@

We originally classified TIC~435903839, with RUWE$_{\rm DR3}$=17.7, as
an ``ambiguous'' CR with a 10.8\,hr period, because this period
minimized the dispersion in the phase-folded light curve.  More
careful inspection revealed an impostor:
this source is a photometric blend of two ordinary rotating stars with
$P_0$=3.60\,hr, and $P_1$=5.41\,hr, giving a beat period $(P_0^{-1} -
P_1^{-1})^{-1}$ of 10.8\,hr.  This is a novel false positive scenario
for CRs: two rapid rotators near the 3:2 period commensurability.
The beat between the two rotation signals produces the apparent CR
signal.  Such false positives can be excluded through careful
accounting of all peaks in a periodogram.  For instance, TIC~435903839
shows a peak at 16.27\,hr, which is not an integer multiple of the
dispersion-minimizing 10.82\,hr period.

@@

While CR periods appear to remain constant to within measurement
precision over thousands of cycles, the light curve shapes evolve over
typical timescales of 10 to 1{,}000 cycles (e.g.
Figures~\ref{fig:evoln} and~\ref{fig:lpriver0}).  The CRs are
therefore quasiperiodic, with quality factors of $Q$$\approx$100.
While this observation is consistent with the analyses by
\citet{2022AJ....163..144G} and \citet{2023ApJ...945..114P}, the fact
that all CR dips seem to be transient over timescales of more than
1{,}000 cycles (Figure~\ref{fig:evoln}) marks a qualitative departure
from the ``persistent'' {\it vs.} ``transient'' flux dip distinction
previously described by \citet{2017AJ....153..152S}.

@@

\subsubsection{What is a prominence?}
The prominence idea is based on a loose analogy with quiescent
prominences/filaments in the solar corona that last as long as a few
weeks \citep[see][]{2015ASSL..415.....V}.  In the context of the Sun,
a prominence is a clump of cold, partially ionized hydrogen viewed in
emission against the dark backdrop of space.  A filament is the same
clump of plasma, but viewed in absorption against the solar disk.  In
an extrasolar context, spectroscopic detections of transient Balmer-
and resonance-line absorption seen for stars such as AB~Dor and
Speedy~Mic
\citep[e.g.][]{1989MNRAS.238..657C,1993MNRAS.262..369J,2006MNRAS.365..530D,2016MNRAS.463..965L}
have been interpreted as prominences that scatter a star's
chromospheric emission \citep[see][]{1989MNRAS.238..657C}.  The
short-term mechanical stability of such gas configurations is
theoretically plausible for rapid rotators
\citep{2000MNRAS.316..647F,2022MNRAS.514.5465W}.   To our best
knowledge, this class of spectroscopic observation also has no viable
alternative explanations.

We performed a simple visual examination of the TESS light curves for
five prominence-hosting systems studied by
\citet{2019MNRAS.482.2853J}--AB~Dor, Speedy~Mic, LQ Lup, HK Aqr, and
V374 Peg--and detected no CR behavior.  While individual prominences
may only last one to tens of rotation cycles, the prominence system
itself is thought to always be ``on'', due to the repeatable
detectability of spectroscopic transients \citep[e.g.][and references
therein]{1990MNRAS.247..415C}.  Assuming that spectroscopically
observable prominence systems indeed do not turn off, this would imply
that they are not always accompanied by photometric CR-like dips: a
link between the spectroscopic prominences that may exist around
rapidly rotating low-mass stars and the CR phenomenon has yet to be
made.

@@

CRs show broadband flux variations that can be 1-2$\times$ deeper in
the blue than in the red
\citep{2017PASJ...69L...2O,2020AJ....160...86B,2022AJ....163..144G,2023MNRAS.518.2921K}.
Dust can naturally explain this chromaticity, since it has a larger
absorption cross-section in the blue than the red
\citep[e.g.][]{1989ApJ...345..245C}.  Gas might also explain the observed
chromaticities \citep{1992oasp.book.....G}.  While bound-bound
absorption can be excluded, since it provides opacity only at narrow
resonant lines, the hydrogen opacity due to bound-free
absorption is ``jagged'' \citep[see][Figure 8.5 and
Eq.~8.8]{1992oasp.book.....G}, such that at temperatures of
$\approx$3{,}000\,K to $\approx$10{,}000\,K the opacity can be larger
at blue wavelengths than at red wavelengths.  Bound-free absorption of
${\rm H}^{-}$ is often important at such temperatures, but this
opacity source is stronger in the red than the blue, the opposite of
what is required to produce deeper dips in the blue than in the red.
Likewise, Thomson scattering is too gray to be the dominant opacity
source.

@@

The observed lifetime of the CR phenomenon could provide another way
to discern between the gas and dust clump scenarios.  Based on the
available statistics from e.g.  \citet{2022AJ....164...80R} and
references therein, it seems plausible that CR occurrence decreases
with stellar age from $\approx$3\% at 10\,Myr (Sco-Cen),
to $\approx$1\% at 100\,Myr (Pleiades),
down to 0\% by the $\approx$700\,Myr
age of Praesepe.  This is odd in the context of the prominence
scenario, because pre-main-sequence M dwarfs spin up over the first
100\,Myr; prominences might therefore be expected to be {\it
more} common at 100\,Myr than at 10\,Myr, under the assumption that the
production of prominences depends only on the stellar rotation rate. 
The dust clump scenario would hold a natural explanation: the lower occurrence of CRs around older stars would
simply reflect a finite supply of dust.
One possible complication however is that the magnetic field topology
of rapidly rotating M dwarfs may depend on factors other than the
rotation rate (e.g.\ the age), which might alter the production of prominences.

@@

In the context of disk-driven planet migration, the stopping location
for the innermost planet is set by the location of the protoplanetary
disk's truncation radius \citep{2018haex.bookE.142I}.  The 
truncation radius is traditionally calculated by equating the magnetic
pressure from the stellar magnetosphere with the ram pressure of the
inflowing gas.  As it happens, the truncation radius is close to the
corotation radius for low accretion rates
\citep[e.g.][]{2016ARA&A..54..135H,2022MNRAS.510.5246L}.  These considerations
invite us to imagine one or more planets migrating inward due to gas
drag, and arriving at $\approx$5-10 stellar
radii before the disk is depleted.

With this picture in mind, it is tempting to attribute features of the
CR light curves to transits of material ejected by planets or
planetesimals.  Young rocky bodies are expected to be hot, and they
might expel either gas or dust.  The Jupiter-Io
system \citep[e.g.][]{2004jpsm.book..537S} is analogous, in that a small rocky
body similarly feeds the construction of a plasma torus.  We emphasize
that while this type of configuration seems a priori plausible, no
direct evidence currently supports it.

The main logical function of the planetesimals would be to serve as a
source for the occulting gas or dust; they would not necessarily need
to explain the observed phases of the observed dips.  The azimuthal
angle of the eventual entrainment could be entirely determined by the
stellar magnetic field.  In this scenario, the obscuring material
would inspiral from one or more rocky bodies well beyond the
corotation radius.  The planetesimals themselves would not necessarily
need to transit.  However if they did, they would need to be $\lesssim
1$\,$R_\oplus$ based on their non-detections in the TESS data.
Possibly analogous systems include K2-22 \citep{2015ApJ...812..112S}
and KOI-2700 \citep{2014ApJ...784...40R}, though the obscuring
material in the CRs would need to be observed much further from the
emitting planet than for those two examples.

@@

Although this outlier is intriguing, stars without infrared excesses
generally have more stable optical light curves than those with
infrared excesses.  While some dippers may evolve into CRs after the
disk is mostly gone, this would be generically expected based on the
population statistics: young objects become old.  There may be no
other causal connection between the two evolutionary stages.  With
that said, a common mystery between the CRs and dippers is how exactly
the {\it narrowness} of the dust-induced dimmings is produced.  It is
not unreasonable to imagine a similar mechanism operating for both
types of object, tied perhaps to a shared magnetic topology, or
perhaps to a preference for dust to inspiral to the star in clumped
structures.

@@

If the disappearance of a dip represents the permanent loss of the
obscuring material -- for example, if it is the result of a dust clump
being accreted or ejected -- then we can also estimate the rate at
which mass is flowing through the structures that lead to dips.  For
instance, LP~12-502 showed three ``state-switch'' events over the six
months of available TESS observations, during cycles 261, 309, and
1241 (Figure~\ref{fig:lpriver0}).  The other source for which we
performed a comparable analysis, TIC~300651846
(Appendix~\ref{app:tic3006}), showed two state-switches over 11
months.  In all such cases, at least one dip turned off.  For purposes
of estimation, we will take LP 12-502 as our prototype.  Assuming the
occulting material is dust, the corresponding $\dot{M} \equiv M\cdot
{\rm d}N/{\rm d}t$ time-averaged over six months is
%$\approx$3$\times$$10^{-18} M_\odot\,{\rm yr}^{-1}
$\approx
1\times10^{-12} M_\oplus\,{\rm yr}^{-1}$.  Considered cumulatively
over the $\approx$$10^8$ years for which the CR phenomenon is
observed, this yields a cumulative moved dust mass of
$10^{-4}\,M_\oplus$, of order the Solar System's asteroid belt.  If
the occulting material is gas, the lower mass bounds would be of order
100 times larger.  For cases in which we observe the {\it growth} of
dips, such as the Sector~29 data for TIC~224283342, or Sector~5 of
TIC~294328885, the dip depths typically increase by of order a few
percent over ten to twenty days.  This growth rate yields a mass flux
one order of magnitude larger than the earlier estimate.

@@

In this work, we searched 120-second cadence TESS data collected from
2018 July to 2022 September for complex rotators (CRs).  The target
stars were \nstarssearched\ late-K and early-to-mid M dwarfs within 150\,pc
and with TESS magnitudes $T$<16. 
The selection function included $>$$80\%$ of such stars within 30\,pc, and $<$$10\%$ of such stars at distances exceeding 100\,pc (Figure~\ref{fig:completeness}).

We found \ngoods\ objects that showed complex
quasiperiodic behavior over at least one TESS sector.  These \ngoods\
bona fide CRs are listed in Table~\ref{tab:thetable}.  This table also
includes \nmaybes\ ambiguous CRs, whose designation is
less certain, and \ndebunked\ impostors.  We inferred ages for all but
two of the \nallcands\ objects based on memberships in young stellar assocations; we
also derived temperatures and radii using SED fitting, and 
inferred stellar masses by interpolating against stellar evolutionary
models.  We caution that our sample is far from being
volume-limited and is not even magnitude-limited:
the TESS 120-second stellar sample had a heterogeneous
selection function which may have been biased in favor of young stars
over field stars.  Previous work however has shown that $\approx$1-3\%
of M dwarfs younger than $\approx$100\,Myr show the CR phenomenon
\citep{2016AJ....152..114R,2022AJ....163..144G}.

@@

      on a quasi-sinusoidal signal that presumably originates from
      inhomogeneities in the stellar surface brightness
      distribution (starspots and faculae).  The
      only viable explanation currently known for sharp dips being superposed on
      the starspot signals is that concentrations (``clumps'')
      of circumstellar material corotate with the star.

@@

      the shortest dips, the clumps must also
      be capable of having sizes comparable to the star.

@@

%{\bf AUTHOR CONTRIBUTION STATEMENTS: PLEASE REVISE IF APPROPRIATE)}
LGB conceived the project, performed the dip-counting search, light curve classification, cluster membership, SED, variability, and secondary-period analyses, and wrote the manuscript.
RJ and SR performed the Fourier-based analysis and contributed to light curve classification.
LR cross-examined the light curve classification, and contributed an independent SED analysis.
AD identified the magnetic B star connection.
LAH contributed to project design and to the interpretation of the light curves.
JNW and SR significantly improved the clarity of the manuscript.
G\'AB acquired and maintained the servers used to run the dip-finding pipeline.
%-----
%{\bf JNW, MNG, GRR to confirm coauthorship \&/or to help out on interpretation!}
% Pending!
%HI and AH designed and maintain the architecture used to acquire and process the HIRES spectra.
All authors assisted in manuscript revision.
